### 2021-04-07 article/sat.tex

\section{About the construction of lower bounds for weak Schur numbers using a computer}

In this section, we first reframe the question of the existence of (weakly) sum-free partitions as a boolean
satisfiability (SAT) problem. We then provide evidence which indicates that the main assumption made by papers which
found the previous best known lower bounds for weak Schur numbers may not be correct. Finally, we obtain stronger
results than those previously known for \(WS(5)\) while gaining several orders of magnitude in computation time by 
giving additional information to the SAT solver without losing in generality. In this section, we assume that the subsets are ordered.

\subsection{Reformulation as a SAT problem}

We encode the existence of (weakly) sum-free partitions as propositional formulae like in \hyperlink{label3}{\textbf{[3]}} and then use 
SAT solvers to determine whether these formulae are satisfiable.

\begin{definition}
	A \textit{literal} is either a variable \(v\) (a positive literal) or the negation \(\bar{v}\) of a variable \(v\) (a negative literal) where \(v\) 
	takes a truth value: \(true\) or \(false\). A \textit{clause} is a disjunction of literals and a \textit{formula} is a conjunction of clauses: it 
	is a propositional formula in \textit{conjonctive normal form} (CNF).
\end{definition}

\begin{definition}
	An \textit{assignment} is a function from a set of variables to the truth values \(true\) (1) and \(false\) (0). A literal \(l\) is 
	\textit{satisfied} (\textit{falsified}) by an assignment \(\alpha\) if l is positive and \(\alpha(var (l)) = 1\) 
	(resp. \(\alpha(var (l)) = 0\)) or if it is negative and \(\alpha(var (l)) = 0\) (resp. \(\alpha(var (l)) = 1\)). A clause is \textit{satisfied} 
	by an assignment \(\alpha\) if it contains a literal that is satisfied by \(\alpha\). Finally, a formula is \textit{satisfied} by an assignment 
	\(\alpha\) if all its clauses are satisfied by \(\alpha\). A formula is \textit{satisfiable} if there exists an assignment that satisfies it; 
	otherwise it is \textit{unsatisfiable}.
\end{definition}

We then encode the existence of a partition of \([\![1,n]\!]\) in \(k\)  weakly sum-free subsets as follows: for every integer 
\(i \in [\![1,n]\!]\), take \(k\) variables \(x^{(i)}_{1}, ..., x^{(i)}_{k}\) and for every \(\forall c \in [\![1,k]\!], x^{(i)}_c = 1 \iff i \in A_c\). 
The corresponding clauses are:

\begin{itemize}
	\item \textbf{sumfree:} \(\forall c \in  [\![1,k]\!], \forall (i, j) \in [\![1,n]\!]^2, (i \neq j ~ \text{and} ~ i + j \leq n) \implies \lnot x^{(i)}_c 
	\lor  \lnot x^{(i)}_c \lor \lnot x^{(i+j)}_c\)
	\item \textbf{union:} \(\forall i \in [\![1,n]\!], x^{(i)}_1 \lor ... \lor x^{(i)}_k\)
	\item \textbf{disjoint:} \(\forall i \in [\![1,n]\!],\forall (c_1, c_2) \in  [\![1,k]\!]^2, c_1 \neq c_2 \implies \neg x^{(i)}_{c_1} \lor \neg x^{(i)}_{c_2}\)
\end{itemize}

In the above formula, every color plays a symmetric role. Hence the search space can reduced by \(k!\) by ordering the subsets, that is by 
enforcing that \(m_1 < ... < m_k\). The corresponding clauses are: \linebreak
\textbf{symmetry breaking:}  \(x^{(1)}_1 = 1\) and \(\forall c \in [\![2,k-1]\!], \forall i \in [\![1,WS(c - 1)+1]\!], x^{(1)}_c \lor ... \lor x^{(i)}_c \lor \neg x^{(i+1)}_{c+1}\)

\begin{remark}
	The performance of
\end{remark}

\subsection{The search space previously used in computer search for lower bounds may not contain the optimal partitions}

Rowley's new lower bound for \(WS(6)\) (642)  \hyperlink{label2}{\textbf{[2]}} was a quite significant improvement upon the former 
best known lower bound (582). This previous lower bound was found using a computer (often with Monte-Carlo) methods and by making the 
assumption that a good partition for \(WS(n+1)\) starts with a good partition for \(WS(n)\) which is true for small values of \(n\).
Therefore, one may wonder whether the limiting factor is the assumption or the methods used to search for partitions. It appears that 
the latter is correct.

\begin{computational theorem}
	There is no weakly sum-free partition of \([\![1,583]\!]\) in 6 parts such that:
	\begin{itemize}
		\item \(m_5 \ge 66\)
		\item \(m_6 \ge 186\)
		\item \([\![210,349]\!] \subset A_6\)
	\end{itemize}  
\end{computational theorem}

This result was obtained in 8 hours with the SAT solver plingeling \hyperlink{label5}{\textbf{[5]}} on a 2.60 GHz Intel i7 processor PC.
However, simply encoding the existence of such a partition as explained in the previous subsection would not result in a reasonable 
computation time. In order to help the SAT solver, we add additional information in the propositional formula. We did not quantify the 
speedup, but it most likely allows us to gain several order of magnitude in computation time as we explain in the next subsection.

\par
Every weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets starts with the following sequence 1121222133. Then 11 is always 
either in subset 1 or 3, 12 is always in subset 3 and so on. For every integer in \([\![1,65]\!]\), we computed in which subset it can appear. 
By using this constraints, we could then compute for every integer in \([\![1,185]\!]\), in which subset it can appear in a weakly sum-free 
partition of  \([\![1,185]\!]\) which starts with a  weakly sum-free partition of \([\![1,65]\!]\) in 4 subsets. Adding these constraints to the 
formula corresponding to the above theorem gives additional information to the SAT solver without losing in generality.

\par
The above theorem shows that the previous lower bound for \(WS(6)\) is optimal in the search space considered by the papers which found it. 
Therefore, finding a partition of \([\![1,n]\!]\) in 6 weakly sum-free subsets for some \(n \geq 590\) which does not have a template-like structure 
would be extremely interesting since it could give indications on a new search space for improving lower bounds with a computer. More generally, 
it questions the search space previously used for finding lower bounds for \(WS(n)\) with a computer. In particular, to our knowledge every paper 
that found the lower bound \(WS(5) \geq 196\) used this assumption [reference necessaire]. Therefore one may wonder if this actually a good lower 
bound. In the next subsection, we give properties that a partition of \([\![1,197]\!]\) in 5 weakly sum-free subsets has to verify.

\subsection{Weak Schur number five}
As explained in the previous subsection, the search space used for showing that  \(WS(5) \geq 196\) may not contain optimal solution. In this subsection, 
we give necessary conditions for a hypothetical partition of \([\![1,197]\!]\) in 5 weakly sum-free subsets using the same type of methods as in the
previous subsection.

\par
First

\begin{computational theorem}
	If there exists a partition of \([\![1,197]\!]\) in 5 weakly sum-free subsets then \(m_5 \leq\). 
\end{computational theorem}

More precisely, we verified the following results using the glucose SAT solver [reference].

\begin{tabular}{| c | *{21}{ p{2mm} |}}
	\hline
	\(m_4\)                   &   4   &   5   &   6  &   7   &   8   &   9   &  10  &  11  &  12  &  13  &  14  &  15  &  16  &  17  &  18  &  19  &  20  &  21  &  22  &  23  &  24  \\ 
	\hline
	\(WS(4 | m_4) + 1\) &  55  &  59  &  60  &  59  &  59  &  60  &  60  &  60  &  60  &  64  &  63  &  64  &  61  &  64  &  63  &  65  &  65  &  65  &  65  &  66  &  67  \\
	\hline
	\(max~m_5\)           &        &       &        &        &       &        &       &        &        &       &        &       &        &       &        &        &       &        &       &        &  53  \\
	\hline
\end{tabular}