### 2021-04-15 article/state_of_the_art.tex
Texte

\section{State of the art}

Before Rowley, the best lower bounds were obtained thanks to a random
tree search. Indeed, the postulate stating that every good weakly sum-
free partition into \(n+1\) colours starts with a weakly sum-free
partition into \(n\) colors leads to a recursive method. However,
these values were improved by Rowley using a theoretical and explicit
construction for special cases. It's interesting to notice that before
that there were no theoretical approach such as Abott and Hanson's
\cite{AbottHanson} for weak Schur.