### 2021-09-23 article/article.tex
affiliations + keywords

\documentclass[3p]{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{MyMnSymbol}

\usepackage[english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{float}

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}

\renewcommand{\qedsymbol}{\(\square\)}

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{computational theorem}[definition]{Computational Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{remark}[definition]{Remark}
\newtheorem{proposition}[definition]{Proposition}

\newcommand{\WS}{\mathit{WS}}
\newcommand{\docarraystretch}{1.2}
\newcommand{\resetarraystretch}{\renewcommand{\arraystretch}{\docarraystretch}}
\resetarraystretch

\title{New lower bounds for Schur and weak Schur numbers}

\author[1]{Romain Ageron\fnref{fn1}}
\ead{romain.ageron@student-cs.fr}

\author[1]{Paul Casteras\fnref{fn1}}
\ead{paul.casteras@student-cs.fr}

\author[1]{Thibaut Pellerin\fnref{fn1}}
\ead{thibaut.pellerin@student-cs.fr}

\author[1]{Yann Portella\fnref{fn1}}
\ead{yann.portella@student-cs.fr}

\author[2]{Arpad Rimmel\fnref{fn2}}
\ead{arpad.rimmel@centralesupelec.fr}

\author[2]{Joanna Tomasik\corref{cor1}\fnref{fn2}}
\ead{joanna.tomasik@centralesupelec.fr}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{These authors contributed equally.}
\fntext[fn2]{These authors supervised the work of the main authors.}

\address[1]{\textit{CentraleSup\'elec}, Universit\'e Paris-Saclay}
\address[2]{LRI, \textit{CentraleSup\'elec}, Universit\'e Paris-Saclay, F-91405, Orsay, France}

\begin{document}

\begin{abstract}

This article provides new lower bounds for both Schur and weak Schur numbers. These results were obtained by 
continuing on Rowley's "template"-based approach for Schur and Ramsey numbers. We also generalized the concept 
of "template" to weak Schur numbers. Finding new templates allows us to get explicit partitions improving lower bounds 
as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). The new lower 
bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), \(\WS (9) \geqslant 22\,536\) and 
\(\WS (10) \geqslant 71\,256 \).

\end{abstract}

\begin{keyword}
Schur number \sep weak Schur number \sep Ramsey theory \sep Sumfree partition
\MSC[2020]{05D10 \sep 05A17 \sep 11B75 \sep 11B81}
\end{keyword}

\maketitle

\input{introduction.tex}

\input{schur.tex}

\input{weakschur.tex}

\section{Conclusions and future work}

\qquad These new results come from an extension of Rowley's template-based approach for Ramsey graphs and 
Schur numbers which is relatively new. Therefore, we will not be surprised if lower bounds are later improved 
using better templates. Moreover, studying specifically \(S^+(n)\) and \(\WS^+(n)\) might be of interest as they
 are closely related to Schur and weak Schur numbers. In order to find new templates, algorithms based on 
randomness such as Monte-Carlo algorithms may prove to be very useful. This could be the subject of a future work.

\section{Acknowledgments}

\qquad We would first like to thank our professors, Joanna Tomasik and Arpad Rimmel, who guided and advised
us during the whole research and writing. We also would like to address warm thanks
to Fred Rowley who kindly answered our many questions and with who it was a pleasure to exchange. 
Last but not least, we would like to thank Marijn Heule for providing the backdoors used in the computation of the 
new lower bound for \(\WS(6)\).

\bibliography{biblio}
\bibliographystyle{elsarticle-num.bst}

\include{appendix.tex}

\end{document}