### 2021-09-23 article/article.tex
abstract + future works + .!!!!

\documentclass[3p]{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{MyMnSymbol}

\usepackage[english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{float}

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}

\renewcommand{\qedsymbol}{\(\square\)}

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{computational theorem}[definition]{Computational Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{remark}[definition]{Remark}
\newtheorem{proposition}[definition]{Proposition}

\newcommand{\WS}{\mathit{WS}}
\newcommand{\docarraystretch}{1.2}
\newcommand{\resetarraystretch}{\renewcommand{\arraystretch}{\docarraystretch}}
\resetarraystretch

\title{New lower bounds for Schur and weak Schur numbers}

\author[1]{Romain Ageron\fnref{fn1}}
\ead{romain.ageron@student-cs.fr}

\author[1]{Paul Casteras\fnref{fn1}}
\ead{paul.casteras@student-cs.fr}

\author[1]{Thibaut Pellerin\fnref{fn1}}
\ead{thibaut.pellerin@student-cs.fr}

\author[1]{Yann Portella\fnref{fn1}}
\ead{yann.portella@student-cs.fr}

\author[2]{Arpad Rimmel\fnref{fn2}}
\ead{arpad.rimmel@centralesupelec.fr}

\author[2]{Joanna Tomasik\corref{cor1}\fnref{fn2}}
\ead{joanna.tomasik@centralesupelec.fr}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{These authors contributed equally.}
\fntext[fn2]{These authors supervised the work of the main authors.}

\address[1]{\textit{CentraleSup\'elec}, Universit\'e Paris-Saclay}
\address[2]{LRI, \textit{CentraleSup\'elec}, Universit\'e Paris-Saclay, F-91405, Orsay, France}

\begin{document}

\begin{abstract}

This article provides new lower bounds for both Schur and weak Schur numbers by exploiting a "template"-based approach. 
The concept of "template" is also generalized to weak Schur numbers. Finding new templates leads to explicit partitions 
improving lower bounds as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). 
The new lower bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), 
\(\WS (9) \geqslant 22\,536\) and \(\WS (10) \geqslant 71\,256 \).

\end{abstract}

\begin{keyword}
Schur number \sep Weak Schur number \sep Ramsey theory \sep Sumfree partition
\MSC[2020]{05D10 \sep 05A17 \sep 11B75 \sep 11B81}
\end{keyword}

\maketitle

\input{introduction.tex}

\input{schur.tex}

\input{weakschur.tex}

\section{Conclusion and future work}

The new results we provide come from an extension of a template-based approach for Ramsey graphs
and Schur numbers. Our theoretical work allowed to improve the lower bounds on Schur and Weak Schur numbers and have
a better understanding of this relatively new approach. Furthermore we introduced the new quantities \(S^+(n)\) and
\(\WS^+(n)\), which may have an interest as they are closely related to the problem at hand.

Given that the introduction of templates is quite recent, we expect the bounds to be improved as this special type of partition 
becomes better understood and larger templates are found. For example, one may try to use
algorithms based on randomness such as Monte-Carlo algorithms to find better templates.

The best lower bound \(\WS(6) \geqslant 583\) achieved with a computer search using Monte-Carlo methods 
\cite{EliahouBook} is significantly lower than those obtained with a template-like structure (\(\WS(6) \geqslant 642\) \cite{RowleyWS}, 
\(\WS(6) \geqslant 646\) in the present article). We have evidence that 583 is the maximal value in the search space considered by 
\cite{EliahouBook,Bouzy2015AnAP,Rafilipojaona}. Therefore, it questions the assumption that large partitions for \(\WS(n+1)\) 
start with large partitions for \(\WS(n)\). It also indicates that \(\WS(5)\) might need further investigation since the current lower 
bound \(\WS(5) \geqslant 196\) \cite{ELIAHOU2012175} it was obtained by considering the same type of search space.

\section{Acknowledgments}

\qquad We would first like to thank our school \textit{CentraleSupélec} and its pole "Formation \`a la recherche" which allowed us to work
on this project during our studies. We also would like to address warm thanks to Fred Rowley who kindly answered our many questions
and with who it was a pleasure to exchange. Last but not least, we would like to thank Marijn Heule for providing the backdoors
used in the computation of the new lower bound for \(\WS(6)\).

\bibliography{biblio}
\bibliographystyle{elsarticle-num.bst}

\include{appendix.tex}

\end{document}