### 2021-09-24 article/article.tex
conclusion

\documentclass[3p]{elsarticle}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{MyMnSymbol}

\usepackage[english]{babel}
\usepackage[hyphens]{url}
\usepackage{hyperref}
\usepackage[utf8]{inputenc}
\usepackage{siunitx}
\usepackage{float}

\usepackage[table]{xcolor}
\usepackage{multirow}
\usepackage{nicematrix}

\renewcommand{\qedsymbol}{\(\square\)}

\newtheorem{definition}{Definition}[section]
\newtheorem{notation}[definition]{Notation}
\newtheorem{theorem}[definition]{Theorem}
\newtheorem{computational theorem}[definition]{Computational Theorem}
\newtheorem{corollary}[definition]{Corollary}
\newtheorem{remark}[definition]{Remark}
\newtheorem{proposition}[definition]{Proposition}

\newcommand{\WS}{\mathit{WS}}
\newcommand{\docarraystretch}{1.2}
\newcommand{\resetarraystretch}{\renewcommand{\arraystretch}{\docarraystretch}}
\resetarraystretch

\title{New lower bounds for Schur and weak Schur numbers}

\author[1]{Romain Ageron\fnref{fn1}}
\ead{romain.ageron@student-cs.fr}

\author[1]{Paul Casteras\fnref{fn1}}
\ead{paul.casteras@student-cs.fr}

\author[1]{Thibaut Pellerin\fnref{fn1}}
\ead{thibaut.pellerin@student-cs.fr}

\author[1]{Yann Portella\fnref{fn1}}
\ead{yann.portella@student-cs.fr}

\author[2]{Arpad Rimmel\fnref{fn2}}
\ead{arpad.rimmel@centralesupelec.fr}

\author[2]{Joanna Tomasik\corref{cor1}\fnref{fn2}}
\ead{joanna.tomasik@centralesupelec.fr}

\cortext[cor1]{Corresponding author}
\fntext[fn1]{These authors contributed equally.}
\fntext[fn2]{These authors supervised the work of the main authors.}

\address[1]{\textit{CentraleSup\'elec}, Universit\'e Paris-Saclay}
\address[2]{LRI, \textit{CentraleSup\'elec}, Universit\'e Paris-Saclay, F-91405, Orsay, France}

\begin{document}

\begin{abstract}

This article provides new lower bounds for both Schur and weak Schur numbers by exploiting a "template"-based approach. 
The concept of "template" is also generalized to weak Schur numbers. Finding new templates leads to explicit partitions 
improving lower bounds as well as the growth rate for Schur numbers, weak Schur numbers and Ramsey numbers \(R_n(3)\). 
The new lower bounds include \(S(9) \geqslant 17\,803\), \(S(10) \geqslant 60\,948\), \(\WS(6) \geqslant 646\), 
\(\WS (9) \geqslant 22\,536\) and \(\WS (10) \geqslant 71\,256 \).

\end{abstract}

\begin{keyword}
Schur number \sep Weak Schur number \sep Ramsey theory \sep Sumfree partition
\MSC[2020]{05D10 \sep 05A17 \sep 11B75 \sep 11B81}
\end{keyword}

\maketitle

\input{introduction.tex}

\input{schur.tex}

\input{weakschur.tex}

\section{Conclusion and future work}

We exhibit new templates for Schur and Ramsey numbers. We also generalize the concept of template to weak Schur numbers. 
These templates provide new general inequalities of the form \(S(n+k) \geqslant a S(n) + b\) and \(\WS(n+k) \geqslant a S(n) + b\) 
as well as new lower bounds for both Schur and weak Schur numbers. We also give a new lower bound \(\WS(6) \geqslant 646\) with 
a method that can yield slight improvements for fixed values of \(n\) over lower bounds obtained with the inequality 
\(\WS(n+1) \geqslant 4 S(n) + 2\).

Given that the introduction of templates is quite recent, we expect the bounds to be improved as this special type of partition 
becomes better understood and larger templates are found. One may try to find better templates by using heuristics or 
approximation algorithms, such as Monte-Carlo algorithms for instance.

The best lower bound \(\WS(6) \geqslant 583\) achieved with a computer search using Monte-Carlo methods 
\cite{EliahouBook} is significantly lower than those obtained with a template-like structure (\(\WS(6) \geqslant 642\) \cite{RowleyWS}, 
\(\WS(6) \geqslant 646\) in the present article). We have evidence that 583 is the maximal value in the search space considered by 
\cite{EliahouBook,Bouzy2015AnAP,Rafilipojaona}. Therefore, it questions the assumption that large partitions for \(\WS(n+1)\) 
start with large partitions for \(\WS(n)\). It also indicates that \(\WS(5)\) might need further investigation since the current lower 
bound \(\WS(5) \geqslant 196\) \cite{ELIAHOU2012175} was obtained by considering the same type of search space.

\section{Acknowledgments}

\qquad First we would like to thank our school \textit{CentraleSupélec} and its pole "Formation \`a la recherche" which allowed us to work
on this project during our studies. We also would like to address warm thanks to Fred Rowley who kindly answered our many questions
and with who it was a pleasure to exchange. Last but not least, we would like to thank Marijn Heule for providing the backdoors
used in the computation of the new lower bound for \(\WS(6)\).

\bibliography{biblio}
\bibliographystyle{elsarticle-num.bst}

\include{appendix.tex}

\end{document}